Coupling PDEs on 3D-1D domains with Lagrange multipliers

### 2019-10-19 lm3d1d_vs1/3d1d_coupled_LM.tex
update

Considering uniform refinements of the initial mesh, Table \ref{tab:error_conform}
lists the errors of formulations \eqref{eq:problem1_simple} and \eqref{eq:problem2_simple}
on the benchmark problem. It can be seen the error in $u$ and $\ud$ in $H^1$ norm
converges linearly (as can be expected due to $P_1$ element discretization).
Moreover, the error of the Lagrange multiplier approximation in $H^{-1/2}$ norm
decreases quadratically. In the light of $P_1$ discretization this rate appears
superconvergent. We speculate that the result is (i) due to the fact that the
exact solution is particularly simple, $L=0$, and (ii) the fact that the error
is interpolated into the same finite element as the approximation $L_h$. We remark
that for $u$ and $\ud$ the error is interpolated into FE space of piecewise
quadratic \emph{discontinous} functions. Evaluating the fractional norm in higher
order discontinous spaces is prohibitively costly. \textbf{Should we try for line formulation
where  this is okay?}
%
\begin{table}
  \scriptsize{
    \begin{minipage}{0.49\textwidth}
  \begin{center}
    \begin{tabular}{c|ccc}
      \hline
    $h$ & $\norm{u-u_h}_{1, \Omega}$ & $\norm{\ud-\udh}_{1, \Lambda}$ & $\norm{L-L_H}_{-1/2,\Gamma}$\\
      \hline
4.3E-1 & 3.4E0(--) & 5.3E-1(--) & 2.9E0(--)\\
2.2E-1 & 1.7E0(0.99) & 2.6E-1(1.06) & 6.1E-1(2.25)\\
1.1E-1 & 8.7E-1(0.99) & 1.3E-1(1.02) & 1.4E-1(2.13)\\
5.4E-2 & 4.4E-1(1.00) & 6.3E-2(1.00) & 3.4E-2(2.03)\\
2.7E-2 & 2.2E-1(1.00) & 3.1E-2(1.00) & 8.6E-3(2.00)\\
\hline
  \end{tabular}
  \end{center}
  \end{minipage}
    }
    \vspace{5pt}
  \scriptsize{%(ii)
    \begin{minipage}{0.49\textwidth}
      \begin{center}
        \begin{tabular}{c|ccc}
      \hline
    $h$ & $\norm{u-u_h}_{1, \Omega}$ & $\norm{\ud-\udh}_{1, \Lambda}$ & $\norm{L-L_H}_{-1/2, \Lambda}$\\
      \hline
4.3E-1 & 3.1E0(--) & 5.4E-1(--) & 4.3E-2(--)\\
2.2E-1 & 1.7E0(0.87) & 2.6E-1(1.06) & 1.1E-2(1.99)\\
1.1E-1 & 8.6E-1(0.96) & 1.3E-1(1.02) & 2.7E-3(2.01)\\
5.4E-2 & 4.4E-1(0.99) & 6.3E-2(1.00) & 6.7E-4(2.00)\\
2.7E-2 & 2.2E-1(1.00) & 3.1E-2(1.00) & 1.7E-4(2.01)\\
\hline
  \end{tabular}
  \end{center}
  \end{minipage}
  }
  \caption{Error convergence of \eqref{eq:problem1_simple} and \eqref{eq:problem2_simple}
    on a benchmark problem \eqref{benchmark}. Continuous linear Lagrange
    elements are used.
  }
  \label{tab:error_conform}
\end{table}

In Table \ref{tab:error_conform} one can observe that the two formulations
yield practically identical approximations of $u$ and $\ud$. However, the solution
cost of the two approaches differs. In Table \ref{tab:cost} we summarize the
size of the linear systems solved at each level of refinement and the time
for the iterative solver to converge. Let as first note that the proposed
preconditioners seem robust with respect to discretization parameter as
the iteration counts are clearly bounded. We then see that the solution
time for \eqref{eq:problem1_simple} is up to 2 times longer compared to
\eqref{eq:problem2_simple}. This is in addition to the higher setup costs
of the preconditioner which in our implementation involve solving an eigenvalue
problem for the fractional Laplacian. Therefore it is advantageous to keep
the multiplier space as small as possible.

\begin{table}
  \scriptsize{
  \begin{center}
    \begin{tabular}{c|cc|ccc|ccc}
      \hline
      $h$ & $\dim{V_h}$ & $\dim{V_{\odot, h}}$ & $\dim{Q^{\Gamma}_h}$ & $\#$ & $T\left[s\right]$ & $\dim{Q^{\Lambda}_h}$ & $\#$ & $T\left[s\right]$\\
      \hline
4.33E-01 &125    &5  &40    &27 &0.03  &5  &9  &0.01\\
2.17E-01 &729    &9  &144   &55 &0.10  &9  &19 &0.02\\
1.08E-01 &4913   &17 &544   &62 &0.25  &17 &36 &0.14\\
5.41E-02 &35937  &33 &2112  &64 &1.97  &33 &42 &1.08\\
2.71E-02 &274625 &65 &8320  &64 &18.01 &65 &36 &8.24\\
\hline
    \end{tabular}
    \end{center}
    }
  \caption{Cost comparison of the two formulations. Number of MinRes iterations
    is denoted by $\#$. Time till convergence of the iterative solver (excluding the setup) is shown
    as $T$.
  }
\label{tab:cost}
\end{table}

>>>>

Considering uniform refinements of the initial mesh, Table \ref{tab:error_conform}
lists the errors of formulations \eqref{eq:problem1_simple} and \eqref{eq:problem2_simple}
on the benchmark problem. It can be seen the error in $u$ and $\ud$ in $H^1$ norm
converges linearly (as can be expected due to $P_1$ element discretization).
Moreover, the error of the Lagrange multiplier approximation in $H^{-1/2}$ norm
decreases quadratically. In the light of $P_1$ discretization this rate appears
superconvergent. We speculate that the result is due to the fact that the
exact solution is particularly simple, $L=0$. In case of the results for
\eqref{eq:problem1_simple} the rate can also be due to the fact that the
error is interpolated into the same finite element as the approximation $L_h$.
We remark that for $u$ and $\ud$ the error is interpolated into FE space of piecewise
quadratic \emph{discontinous} functions. For \eqref{eq:problem2_simple} we
evaluate the fractional norm and interpolate the error using piecewise continuous
cubic functions. Evaluating the fractional norm in higher order spaces
for formulation with the multiplier space on $\Gamma$ is prohibitively costly.

%
\begin{table}
  \scriptsize{
    \begin{minipage}{0.49\textwidth}
  \begin{center}
    \begin{tabular}{c|ccc}
      \hline
    $h$ & $\norm{u-u_h}_{1, \Omega}$ & $\norm{\ud-\udh}_{1, \Lambda}$ & $\norm{L-L_H}_{-1/2,\Gamma}$\\
      \hline
4.3E-1 & 3.4E0(--) & 5.3E-1(--) & 2.9E0(--)\\
2.2E-1 & 1.7E0(0.99) & 2.6E-1(1.06) & 6.1E-1(2.25)\\
1.1E-1 & 8.7E-1(0.99) & 1.3E-1(1.02) & 1.4E-1(2.13)\\
5.4E-2 & 4.4E-1(1.00) & 6.3E-2(1.00) & 3.4E-2(2.03)\\
2.7E-2 & 2.2E-1(1.00) & 3.1E-2(1.00) & 8.6E-3(2.00)\\
\hline
  \end{tabular}
  \end{center}
  \end{minipage}
    }
    \vspace{5pt}
  \scriptsize{%(ii)
    \begin{minipage}{0.49\textwidth}
      \begin{center}
        \begin{tabular}{c|ccc}
      \hline
    $h$ & $\norm{u-u_h}_{1, \Omega}$ & $\norm{\ud-\udh}_{1, \Lambda}$ & $\norm{L-L_H}_{-1/2, \Lambda}$\\
      \hline
4.3E-1 & 3.1E0(--)    & 5.4E-1(--)   & 4.4E-2(--)   \\
2.2E-1 & 1.7E0(0.87)  & 2.6E-1(1.06) & 1.1E-2(2.01) \\
1.1E-1 & 8.6E-1(0.96) & 1.3E-1(1.02) & 2.7E-3(2.01) \\
5.4E-2 & 4.4E-1(0.99) & 6.3E-2(1.00) & 6.7E-4(2.01) \\
2.7E-2 & 2.2E-1(1.00) & 3.1E-2(1.00) & 1.7E-4(2.01) \\
1.4E-2 & 1.1E-1(1.00) & 1.6E-2(1.00) & 4.1E-5(2.01) \\
\hline
  \end{tabular}
  \end{center}
  \end{minipage}
  }
  \caption{Error convergence of \eqref{eq:problem1_simple} and \eqref{eq:problem2_simple}
    on a benchmark problem \eqref{benchmark}. Continuous linear Lagrange
    elements are used.
  }
  \label{tab:error_conform}
\end{table}

In Table \ref{tab:error_conform} one can observe that the two formulations
yield practically identical approximations of $u$ and $\ud$. However, the solution
cost of the two approaches differs. In Table \ref{tab:cost} we summarize the
size of the linear systems solved at each level of refinement and the time
for the iterative solver to converge. Let as first note that the proposed
preconditioners seem robust with respect to discretization parameter as
the iteration counts are clearly bounded. We then see that the solution
time for \eqref{eq:problem1_simple} is about 2 times longer compared to
\eqref{eq:problem2_simple}. This is in addition to the higher setup costs
of the preconditioner which in our implementation involve solving an eigenvalue
problem for the fractional Laplacian. Therefore it is advantageous to keep
the multiplier space as small as possible. Let us remark that the missing
results for \eqref{eq:problem1_simple} in Table \ref{tab:cost} and \ref{tab:error_conform}
are due to the memory limitations which we encouter when solving the eigenvalue problem
for Laplacian, which for $h=1.35E-02$ includes cca 32 thousand eigenvalues.
%
\begin{table}
  \scriptsize{
  \begin{center}
    \begin{tabular}{c|cc|ccc|ccc}
      \hline
      $h$ & $\dim{V_h}$ & $\dim{V_{\odot, h}}$ & $\dim{Q^{\Gamma}_h}$ & $\#$ & $T\left[s\right]$ & $\dim{Q^{\Lambda}_h}$ & $\#$ & $T\left[s\right]$\\
      \hline
4.33E-01 &125    &5  &40    &27 &0.03  &5  &9  &0.01\\
2.17E-01 &729    &9  &144   &55 &0.10  &9  &19 &0.02\\
1.08E-01 &4913   &17 &544   &62 &0.25  &17 &36 &0.14\\
5.41E-02 &35937  &33 &2112  &64 &1.97  &33 &42 &1.08\\
2.71E-02 &274625 &65 &8320  &64 &18.01 &65 &36 &8.24\\
1.35E-02 &--     &-- &--    &-- &--    &129 &31 &61.37\\
\hline
    \end{tabular}
    \end{center}
    }
  \caption{Cost comparison of the two formulations. Number of MinRes iterations
    is denoted by $\#$. Time till convergence of the iterative solver (excluding the setup) is shown
    as $T$.
  }
\label{tab:cost}
\end{table}
